\Title

### 2021-01-06 manuscript/manuscript.tex
Review text for horizontal positioning of sources

The ideal number of sources and their locations, both horizontally and
vertically, has been debated since the inception of the equivalent source
technique with \citet{dampney1969}.
The choices made regarding these parameters can play an important role on the
accuracy of the predictions and the computational resources needed to estimate
the coefficients.
An ideal distribution of sources should simultaneously be able to reproduce the
measured data on the survey points and also be able to make accurate
predictions on non-observed locations.

A large number of sources evenly distributed along the survey region are
capable of reproducing the observed data. Nevertheless, the computational load
can be prohibitive and such underdetermined problems are prone to overfiting
the data, leading to poor performance when making predictions. On the other
hand, using very few sources will undoubtedly reduce the computational
requirements but the equivalent source model could be incapable of reproducing
the measured data, leading to inaccurate predictions.

Particular survey characteristics also play a role in the choice of source
distribution.
In a ground survey, observation points are usually located along irregular
paths as well as some possible scattered points.
The coverage of the survey region is usually uneven, often leaving large areas
without any observation.
On the other hand, observations from airborne surveys are located along almost
straight flight lines.
Measurements are usually taken at
a high temporal frequency, leading to the observation points along the flight
lines that are several times closer than the flight line spacing.
This creates an anisotropy in the sampling, which can cause aliasing artifacts
in gridded products.

\subsubsection{Horizontal layout}

The most widely used layouts of equivalent sources are:

\begin{itemize}
  \item
    \emph{Regular grid}: a homogeneous distribution of point sources below the
    survey region at a constant depth. A padding region is often added to help
    reduce edge effect. In practice, often leads to underdetermined problems
    ($N<M$).
  \item
    \emph{Sources below data}: one equivalent source is placed below each
    observation point at a given depth (often dependent on the distance to
    neighboring data points). Therefore, the number of sources is be equal to
    the number of observations ($N=M$).
\end{itemize}

For ground surveys, the \emph{regular-grid} layout needs a sufficiently
small grid spacing to be able to fit the observed data.
This creates an unnecessarily large number of sources in areas where no
observations exist.
In contrast, the \emph{sources-below-data} layout is more likely to accurately
fit the observed data with many fewer sources, reducing the computational load.

For airborne surveys, the \emph{source-below-data} layout may
place an undesirably large number of sources along the flight paths.
This could lead to aliasing effects on the predicted values, such as the
stripes parallel to flight lines that are often observed when griding airborne
magnetic data.
In contrast, the \emph{regular-grid} layout can avoid this effect by evenly
distributing sources and using a continuous source layer (e.g.,
right-rectangular prisms or tesseroids).

We propose a new source layout that could simultaneously reduce the
computational load and remove some of the drawbacks of the regular-grid and
source-below-data layouts:

\begin{itemize}
  \item
    \emph{Block-averaged sources}: point sources are placed in the average
    position of data points that fall within specified spatial blocks
    (Fig.~\ref{fig:block-averaged-sources}). In
    short, this is done by:
    \begin{enumerate}
        \item Dividing the survey region in blocks of equal size
            (Fig.~\ref{fig:block-averaged-sources}a-b).
        \item Computing the median position of the observation points that fall
          inside each block. Blocks without any observation point are ignored.
        \item Assign one point source to the computed median position, shifted
          to a specified depth (Fig.~\ref{fig:block-averaged-sources}c).
    \end{enumerate}
\end{itemize}

The number of sources created by this new layout will be less than or equal to
(if the blocks contain a single data point)
the number of observations.
The overdetermined problems that arise from this layout have a lower
computational load and are less prone to overfiting since the complexity of the
model is lower.
Moreover, the block averaging process can balance the spacing between sources
along a flight line and between adjacent lines.

>>>>

The ideal number of sources and their locations, both horizontally and
vertically, has been debated since the inception of the equivalent sources
technique with \citet{dampney1969}.
The choices made regarding these parameters can play an important role on the
accuracy of the predictions and the computational resources needed to estimate
the source coefficients.
An ideal distribution of sources should simultaneously be able to reproduce the
measured data on the survey points, make accurate predictions on non-surveyed
locations, and minimize the required computational resources.

A large number of evenly distributed sources along the survey region are
capable of reproducing the observed data.
Nevertheless, the computational load can be prohibitive and such
underdetermined problems are prone to overfiting the data, leading to poor
predictive power when interpolating and extrapolating.
On the other hand, using few sources will reduce the computational requirements
but the model may be incapable of reproducing the full spectral content of the
measured data.

Particular survey characteristics also play a role in the choice of equivalent
source distribution.
In a ground survey, observations are usually located along irregular paths and
scattered points.
The coverage of the survey region is often uneven, leaving large areas without
any observation.
On the other hand, observations from airborne surveys are located along almost
straight and closely spaced flight lines.
Measurements are usually taken at a high temporal frequency, leading to the
observation points along the flight lines that are several times closer to each
other than the flight line spacing.
This creates a bias in the sampling, which can cause aliasing artifacts in
gridded products.

\subsubsection{Horizontal source layouts}

The most widely used layouts for distributing equivalent sources horizontally
are:

\begin{enumerate}
  \item
    \emph{Regular grid}: a homogeneous distribution of point sources below the
    survey region at a constant depth. A padding region is often added to help
    reduce edge effects. In practice, it often leads to underdetermined problems
    since a large number of sources is required ($N<M$).
  \item
    \emph{Sources below data points}: one equivalent source is placed below each
    observation point at a given depth (often dependent on the distance to
    neighboring data points). Therefore, the number of sources is be equal to
    the number of observations ($N=M$).
\end{enumerate}

For ground surveys, the \emph{regular-grid} layout needs a sufficiently
small grid spacing to be able to fit the observed data.
This creates an unnecessarily large number of sources in areas where no
observations exist.
In contrast, the \emph{sources-below-data} layout is more likely to accurately
fit the observed data with many fewer sources, reducing the computational load.
But when applied to airborne surveys, the \emph{source-below-data} layout may
place an undesirably large number of sources along the flight paths.
This could lead to aliasing effects on the predicted values, such as the
stripes parallel to flight lines that are often observed when griding airborne
magnetic data.
In contrast, the \emph{regular-grid} layout can avoid this effect by evenly
distributing sources and using a continuous source layer (e.g.,
right-rectangular prisms or tesseroids).

We propose a new way of distributing equivalent sources horizontally that could
simultaneously reduce the computational load and mitigate some of the drawbacks
of existing layouts.
In the \emph{block-averaged sources} layout,
point sources are placed in the average
position of data points that fall within specified spatial blocks
(Fig.~\ref{fig:block-averaged-sources}).
This is done by:

\begin{enumerate}
    \item Dividing the survey region into rectangular blocks of equal size.
    \item Computing the median horizontal position of the observation points
      that fall inside each block. Blocks without any observation point are
      omitted.
    \item Assign one point source to the median horizontal position calculated
      in step 1.
\end{enumerate}

The number of sources created by this new layout will be less than the number
of observations if the block size is chosen appropriately (i.e., making sure
that blocks are large enough to contain more than a single data point).
The overdetermined problem that arises from this layout has a lower
computational load and is less prone to overfiting the data since the model
complexity is lower.
Moreover, the block averaging process can balance the spacing between sources
along a flight line and between adjacent lines, helping to reduce aliasing
effects in generated grids.
